\documentclass[hyperref={colorlinks=true,urlcolor=blue}]{beamer}

\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{hyperref}
\usepackage[english]{babel}
\usepackage[babel=true]{microtype}
\usepackage{color}
\usepackage{listings}

\definecolor{verbgray}{gray}{0.9}

\lstnewenvironment{code}{%
  \lstset{backgroundcolor=\color{verbgray},
  frame=single,
  framerule=0pt,
  basicstyle=\ttfamily,
  columns=fullflexible}}{}

\title{A brief introduction to\\ scalaz-stream}
\author{Frank S. Thomas}
\date{September 18, 2013}

\begin{document}
  \frame{\titlepage}

\begin{frame}{What is it?}
  scalaz-stream is
  \begin{itemize}
    \item a library for stream processing and incremental I/O
    \item resource safe
    \item composable
    %\item expressive
    \item functional: stream transformations are pure values
  \end{itemize}
\end{frame}
 
\begin{frame}[fragile]{Imperative I/O}
  \begin{code}
def linesGt40k1(filename: String): Boolean = {
  val src = scala.io.Source.fromFile(filename)
  try {
    var count = 0
    val lines: Iterator[String] = src.getLines
    while (count <= 40000 && lines.hasNext) {
      lines.next
      count += 1
    }
    count > 40000
  }
  finally src.close
}      
  \end{code}
\end{frame}

\begin{frame}[fragile]{I/0 with \texttt{scalaz-stream}}
  \begin{code}
def linesGt40k2(filename: String): Boolean =
  scalaz.stream.io.linesR(filename).
    drop(40000).
    once.
    runLast.
    run.
    nonEmpty
  \end{code}
\end{frame}

\begin{frame}[fragile]{Stream transformations with \texttt{Process}}
  \begin{code}
trait Process[I,O]

case class Emit[I,O](
    head: Seq[O],
    tail: Process[I,O] = Halt[I,O]())
  extends Process[I,O]

case class Await[I,O](
    recv: I => Process[I,O],
    finalizer: Process[I,O] = Halt[I,O]())
  extends Process[I,O]

case class Halt[I,O]() extends Process[I,O]
  \end{code}
  
  \begin{itemize}
   \item \texttt{Process} is a state machine with three possible states
   \item these must be interpreted by a driver for effects to occur
  \end{itemize}  
\end{frame}

\begin{frame}{Combinators}
  Modify and compose processes:
  \begin{itemize}
    \item \texttt{def map[O2](f: O => O2): Process[I,O2]}
    \item \texttt{def flatMap[O2](f: O => Process[I,O2]): Process[I,O2]}
    \item \texttt{def append(p: => Process[I,O]): Process[I,O]}
    \item \texttt{def filter(f: I => Boolean): Process[I,I]}
    \item \texttt{def take(n: Int): Process[I,I]}
    \item \texttt{def takeWhile(f: I => Boolean): Process[I,I]}
    \item ... and many more
  \end{itemize}
  
  Feed the output of \texttt{this} to the input of \texttt{p2}: 
  \begin{itemize}
    \item \texttt{def pipe[O2](p2: Process[O,O2]): Process[I,O2]}
  \end{itemize}
\end{frame}

\begin{frame}[fragile]{Another Example}
  \begin{code}
import scalaz.stream._
import scalaz.concurrent.Task

val converter: Task[Unit] =
  io.linesR("testdata/fahrenheit.txt").
     filter(s =>
       !s.trim.isEmpty && !s.startsWith("//")).
     map(line =>
       fahrenheitToCelsius(line.toDouble).toString).
     intersperse("\n").
     pipe(process1.utf8Encode).
     to(io.fileChunkW("testdata/celsius.txt")).
     run

val u: Unit = converter.run 
  \end{code}
\end{frame}

\begin{frame}{References}
  \begin{itemize}
   \item \url{https://github.com/scalaz/scalaz-stream}
   \item Chapter 15 of Functional Programming in Scala
     \url{http://manning.com/bjarnason/}
   \item Advanced Stream Processing in Scala
     \url{http://www.youtube.com/watch?v=8fC2V9HX_m8}
   %\item \url{http://stackoverflow.com/questions/tagged/scalaz-stream}
  \end{itemize}
\end{frame}

\end{document}
